Dynamics of Colombo's Top: Tidal Dissipation and Resonance Capture

### 2021-05-29 writeups/Su_weak_tides.tex
added some more plots, send dong draft

In this section, we briefly introduce the spin dynamics of the planet (for a
more detailed treatment, see Paper I). We consider a star of mass $M_\star$
hosting an inner oblate planet of mass $m$ and radius $R$ on a circular orbit
with semi-major axis $a$ and an outer perturber of mass $m_{\rm p}$ on a
circular orbit with semi-major axis $a_{\rm p}$. We assume that the two orbits
are mutually inclined by the angle $I$. Denote $\bm{S}$ the spin angular
momentum and $\bm{L}$ the orbital angular momentum of the planet, and
$\bm{L}_{\rm p}$ the angular momentum of the perturber. The corresponding unit
vectors are $\uv{s} \equiv \bm{S} / S$, $\uv{l} \equiv \bm{L} / L$, and
$\uv{l}_{\rm p} \equiv \bm{L}_{\rm p} / L_{\rm p}$. The spin axis $\uv{s}$ of
the planet tends to precess around its orbital (angular momentum) axis $\uv{l}$,
driven by the gravitational torque from the host star acting on the planet's
rotational bulge. On the other hand, $\uv{l}$ and the disk axis $\uv{l}_{\rm p}$
precess around each other due to gravitational interactions. We assume $S \ll L
\ll L_{\rm p}$, so $\uv{l}_{\rm p}$ and $\uv{l}$ are nearly constant. The
equations of motion for $\uv{s}$ and $\uv{l}$ in this limit are
(\citealp{anderson2018teeter}, Paper I)
\begin{align}
    \rd{\uv{s}}{t}
        &= \omega_{\rm sl}\p{\uv{s} \cdot \uv{l}}\p{\uv{s} \times \uv{l}}
        \equiv \alpha\p{\uv{s} \cdot \uv{l}}\p{\uv{s} \times
        \uv{l}},\label{eq:dsdt1}\\
    \rd{\uv{l}}{t} &= \omega_{\rm lp}\p{\uv{l} \cdot \uv{l}_{\rm p}}\p{\uv{l}
        \times \uv{l}_{\rm p}} \equiv -g\p{\uv{l} \times \uv{l}_{\rm p}},
        \label{eq:dldt1}
\end{align}
where
\begin{align}
    \omega_{\rm sl} &\equiv \frac{3GJ_2 mR^2 M_\star}{2a^3 I\Omega_{\rm s}}
        = \frac{3k_q}{2k}\frac{M_\star}{m}\p{\frac{R}{a}}^3 \Omega_{\rm s},
            \label{eq:wsl}\\
    \omega_{\rm lp} &= \frac{3m_{\rm p}}{4M_\star}\p{\frac{a}{a_{\rm p}}}^3
        n.\label{eq:wlp}
\end{align}
In Eq.~\eqref{eq:wsl}, $\Omega_{\rm s}$ is the spin frequency of the inner
planet, $\mathcal{I} = k mR^2$ (with $k$ a constant) is its moment of inertia
and $J_2 = k_{\rm q}\Omega_{\rm s}^2 (R^3/Gm)$ (with $k_{q}$ a constant) is its
rotation-induced (dimensionless) quadrupole moment [for a body with uniform
density, $k=0.4, k_{\rm q}=0.5$; for rocky planets, $k\simeq 0.2$ and $k_{\rm
q}\simeq 0.17$ \citep[e.g.][]{lainey2016quantification}]. In other studies,
$3k_{\rm q} / 2 k$ is often notated as $k_2 / 2C$
\citep[e.g.][]{millholland_disk}. In Eq.~\eqref{eq:wlp}, $n \equiv
\sqrt{GM_\star/a^3}$ is the inner planet's orbital mean motion,  and we have
assumed $a_{\rm p}\gg a$ and included only the leading-order (quadrupole)
interaction between the inner planet and perturber. Following standard notation,
we define $\alpha = \omega_{\rm sl}$ and $g \equiv -\omega_{\rm 1p} \cos I$
\citep[e.g.][]{colombo1966}.

@@

Using this result, we can then identify the long-term equilibria of the system
when $\Omega_{\rm s}$ is also evolving: these equilibria of the system must both
satisfy $\dot{\Omega}_{\rm s} = 0$ and be a CS that is stable in the presence of
tidal dissipation (i.e.\ satisfying $\rdil{\uv{s}}{t} = 0$).
Figure~\ref{fig:6equilssingle} describes the evolution of
Eqs.~(\ref{eq:dsdt_tide}--\ref{eq:dWsdt_tide}) qualitatively in the coordinates
$(\Omega_{\rm s}, \theta)$, along with the locations of CS1 and CS2. The circled
points in the panels of Fig.~\ref{fig:6equils} satisfy the criteria to be
long-term equilibria, and we call them \emph{tidal Cassini Equilibria} (tCE). We
number tCE1 and tCE2 the tCE that are in CSs 1 and 2 respectively. The
obliquities of the tCE depend on the system architecture, which can be
quantified using the parameter
\begin{align}
    \eta_{\rm sync} &\equiv \big(\eta\big)_{\Omega_{\rm s} = n}
        = \eta \frac{\Omega_{\rm s}}{n}\nonumber\\
        &= \frac{1}{2}\frac{k}{k_{\rm q}}
            \frac{m_{\rm p}m}{M_\star^2}
            \p{\frac{a}{a_{\rm p}}}^3 \p{\frac{a}{R}}^3\cos(I).
            \label{eq:def_etasync}
\end{align}
In Fig.~\ref{fig:6equilssingle}, $\eta_{\rm sync} = 0.06$; the two panels of
Fig.~\ref{fig:6equils} illustrate the change to the tCE when $\eta_{\rm sync} =
0.5$ and $0.7$ respectively.

@@

    \includegraphics[width=\textwidth]{../initial/1_weaktide/6equilssingle.png}

@@

    used $\abs{g}t_{\rm s} = 10^2$). The phase space evolution of the two
    thicker evolutionary trajectories (cyan and pink; those beginning at
    $\theta_{\rm i} = 120^\circ$) are shown in

@@

    }\label{fig:6equilssingle}

@@

    \includegraphics[width=\columnwidth]{../initial/1_weaktide/6equils.png}
    \caption{Same as Fig.~\ref{fig:6equilssingle} but for $\eta_{\rm sync} =
    0.5$ and $\eta_{\rm sync} = 0.7$ in the two panels (labeled). Note in the
    bottom panel that $\eta_{\rm sync} = 0.7 > \eta_{\rm c}$ and tCE1 does not
    exist. In both panels, the light-colored crosses and lines correspond to
    evolutionary trajectories using the same initial conditions as those shown
    in Fig.~\ref{fig:6equilssingle}. The phase space evolution of the thick
    purple trajectory in the bottom panel (starting at $\theta_{\rm i} \approx
    10^\circ$) is shown in Fig.~\ref{fig:trajs3}. }\label{fig:6equils}

@@

    Fig.~\ref{fig:6equilssingle} colored in pink, for which $\eta_{\rm sync} =
    0.06$ and $I = 20^\circ$. In the
    left-most panel, the trajectory's evolution in $\p{\Omega_{\rm s}, \theta}$
    coordinates along with the curves indicating CS1, 2, and $\dot{\Omega}_{\rm
    s} = 0$ are re-displayed from Fig.~\ref{fig:6equilssingle}. The vertical
    dashed lines denote the values of $\Omega_{\rm s}$ for which the phase space
    evolution of the system is displayed in the right four panels. In each of
    these right four panels, the trajectory's evolution for a single
    circulation/libration cycle is displayed in $\p{\cos \theta, \phi}$ phase
    space for the labeled value of $\Omega_{\rm s}$. The system encounters the
    separatrix, undergoes a III $\to$ I transition, and converges to tCE1.

@@

    \caption{Same as Fig.~\ref{fig:trajs1} but for the trajectory in
    Fig.~\ref{fig:6equilssingle} colored in cyan, for which $\eta_{\rm sync} =
    0.06$ and $I = 20^\circ$. The system encounters the
    separatrix, undergoes a III $\to$ II transition, and converges to tCE2. Note
    that the initial condition of this system and that displayed in
    Fig.~\ref{fig:trajs1} have the same initial $\theta_{\rm i}$ and
    $\Omega_{\rm s, i}$ and only different precessional phases $\phi_{\rm i}$.

@@

    \caption{Same as Fig.~\ref{fig:trajs1} but for the purple trajectory
    shown in the bottom panel of Fig.~\ref{fig:6equils}, for which $\eta_{\rm
    sync} = 0.7$ and $I = 20^\circ$. Here, the system
    evolves along CS1 until the separatrix disappears, upon which it experiences
    large obliquity variations that damp due to tidal dissipation. The system
    finally converges to tCE2, the only tCE that exists for this value of
    $\eta_{\rm sync}$ (see bottom panel of Fig.~\ref{fig:6equils}).}\label{fig:trajs3}

@@

There are two important conditions that can change the existence and stability
of the tCE\@. First, if $\eta_{\rm sync} > \eta_{\rm c}$, then tCE1 will not
exist, as can be seen in the bottom panel of
Fig.~\ref{fig:6equils}\footnote{Strictly speaking, $\eta_{\rm sync}$ can be
slightly smaller than $\eta_{\rm c}$, as the planet's spin is slightly
subsynchronous at tCE1 when $\eta \approx \eta_{\rm c}$, see the top panel of
Fig.~\ref{fig:6equils}.}. Secondly, tCE2 may not be stable if the tidal phase
shift is too large. Applying the results of Section~\ref{ss:tidal_equils}, we
find that tCE2 is stable as long as $t_{\rm s} \geq t_{\rm s, c}$ where
\begin{equation}
    t_{\rm s, c} \equiv \frac{\sin \theta_{\rm tCE2}}{\abs{g} \sin
            I}\p{\frac{2n}{\Omega_{\rm s, tCE2}} - \cos \theta_{\rm tCE2}}.
\end{equation}
The tCE2 subscripts denote evaluation at tCE2. When $\eta_{\rm sync} \ll
\eta_{\rm c}$, we can use Eq.~\eqref{eq:def_tce2_approx} to further simplify:
\begin{equation}
    t_{\rm s, c}\approx \frac{1}{\abs{g}\sin I}\sqrt{\frac{2}{\eta_{\rm sync}
        \cos I}}. \label{eq:def_ts_crit}
\end{equation}
Note that this result differs in form from that given in
\citet{fabrycky_otides}, which assume that the orbital precession of $\uv{l}$ is
driven by the central star. Our expression is simpler and is applicable to
systems where the orbital precession of the inner planet is dominated by an
exterior planet.

@@

With this result, we can now consider the final fate of the inner planet's spin.
We begin by qualitatively examining the example trajectories shown in
Figs.~\ref{fig:6equilssingle}, for which we have integrated
Eqs.~(\ref{eq:dsdt_rot},~\ref{eq:dsdt_tide}--\ref{eq:dWsdt_tide}). We discuss
each of the six trajectories in turn:
\begin{itemize}
    \item The trajectory with initial conditions $\Omega_{\rm s, i} = 2.5n$ and
        $\theta_{\rm i} = 10^\circ$ (purple) has an initially prograde spin
        (i.e.\ in zone I, see FIg.~\ref{fig:1contours}) and directly evolves to
        tCE1 as the obliquity damps to zero.

@@

    \item We have chosen two trajectories, both with initial conditions
        $\Omega_{\rm s, i} = 2.5n$ and $\theta_{\rm i} = 120^\circ$ but with
        different initial precessional phases $\phi_{\rm i}$; we first discuss
        the pink trajectory. This trajectory originates in zone III, evolves
        towards the separatrix as tidal friction damps the obliquity, and
        crosses the resonance without being captured, upon which the obliquity
        continues to damp until the system converges to tCE1. The detailed phase
        space evolution of this trajectory is shown in Fig.~\ref{fig:trajs1},
        where the outcome of the separatrix encounter is very visible.

    \item The other trajectory with initial conditions $\Omega_{\rm s, i} = 2.5n$
        and $\theta_{\rm i} = 120^\circ$ (light blue) only differs from the pink
        trajectory in its value of $\phi_{\rm i}$. Nevertheless, this trajectory
        originates in zone III, encounters the separatrix and is captured into
        the resonance (zone II), upon which tidal friction drives the system
        towards tCE2. The detailed phase space evolution of this trajectory is
        shown in Fig.~\ref{fig:trajs2}, where the resonance capture is
        displayed.

@@

    \item The trajectory with initial conditions $\Omega_{\rm s, i} = 0.1n$ and
        $\theta_{\rm i} = 100^\circ$ (teal) also has its obliquity rapidly
        damped to zero as it approaches spin-orbit synchronization and tCE1.
        However, we remark that this initial condition converges to tCE2 when a
        tidal dissipation rate of $\abs{g}t_{\rm s} = 10^3$ is used.
\end{itemize}
For each of these six initial conditions, we also display their evolutionary
trajectories for $\eta_{\rm sync} = 0.5$ and $\eta_{\rm sync} = 0.7$ in the two
panels of Fig.~\ref{fig:6equils}. Of particular interest is the purple initial
condition ($\Omega_{\rm s, i} = 2.5n$ and $\theta_{\rm i} = 10^\circ$) when
$\eta_{\rm sync} = 0.7$. Figure~\ref{fig:trajs3} shows the detailed phase space
evolution of this system, where it can be seen that the system initially evolves
along the stable CS1, but is ejected when $\Omega_{\rm s}$ is sufficiently small
that CS1 ceases to exist, upon which large obliquity variations eventually lead
to convergence to tCE2, the only tCE that exists.

It can be seen that the dynamics are complex and can vary greatly depending on
the parameters used. For tractability, we restrict our attention to initial
conditions satisfying $\Omega_{\rm s, i} \gg n$, and we adopt the fiducial value
$\Omega_{\rm s, i} = 10n$. Then, for many initial $\theta_0$ and $\phi_0$, we
can perform the same integration as above and verify which tCE a given initial
condition evolves into. In Fig.~\ref{fig:Hhists_0_06}, we show the final outcome
for many randomly chosen $\theta_0$ and $\phi_0$ for $\eta_{\rm sync} = 0.06$
and $I = 20^\circ$. We see that the behavior seen in the example trajectories in
Fig.~\ref{fig:6equilssingle} are general: tCE1 is generally reached for spins
initially in zone I (like the purple trajectory in
Fig.~\ref{fig:6equilssingle}), tCE2 is generally reached for spins initially in
zone II (like the red trajectory in Fig.~\ref{fig:6equilssingle}), and a
probabilistic outcome is observed for spins initially in zone III (like the
light blue and pink trajectories in Fig.~\ref{fig:6equilssingle}).
Figures~\ref{fig:Hhists_0_20} and~\ref{fig:Hhists_0_50} show the same results
but for $\eta_{\rm sync} = 0.2$ and $\eta_{\rm sync} = 0.5$. As $\eta_{\rm
sync}$ is increased, more initial conditions reach tCE2. This is both because
there are more systems initially in zone II and because systems initially in
zone III have a higher probability of executing a III $\to$ II transition upon
separatrix encounter. Note also that in Fig.~\ref{fig:Hhists_0_50}, even initial
conditions in zone I are able to reach tCE2; we comment on the origin of this
behavior in the next section.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5Hhists0_06_20.png}
    \caption{\emph{Left:} Each dot indicates which tCE a given initial condition
    $(\theta_{\rm i}, \phi_{\rm i})$ evolve towards (labeled in legend), for
    $\eta_{\rm sync} = 0.06$ and $I = 20^\circ$. The planet's initial spin is
    chosen to be $\Omega_{\rm s, i} = 10n$, and the three zones of phase space
    (see Fig.~\ref{fig:1contours}) for the initial conditions are labeled,
    separated by the separatrix (black line). Note that initial conditions in
    zone I evolve towards tCE1, initial conditions in zone II evolve towards
    tCE2, and initial conditions in zone III have an effectively probabilistic
    outcome. \emph{Right:} Histogram of which tCE a given initial obliquity
    $\theta_{\rm i}$ evolves towards, averaged over $\phi_{\rm
    i}$.
    \textcolor{red}{NB\@: the excess of tCE2 initial conditions at $\cos
    \theta_{\rm i} \approx -1$ is likely an artifact of loss of numerical
    precision, I am working on fixing this.}
    }\label{fig:Hhists_0_06}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5Hhists0_20_20.png}
    \caption{Same as Fig.~\ref{fig:Hhists_0_06} but for $\eta_{\rm sync} =
    0.2$.
    \textcolor{red}{NB\@: the excess of tCE2 initial conditions at $\cos
    \theta_{\rm i} \approx -1$ is likely an artifact of loss of numerical
    precision, I am working on fixing this.}
    }\label{fig:Hhists_0_20}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5Hhists0_50_20.png}
    \caption{Same as Fig.~\ref{fig:Hhists_0_06} but for $\eta_{\rm sync} =
    0.5$. Note that even points above the separatrix can evolve towards tCE2
    here.}\label{fig:Hhists_0_50}
\end{figure}

>>>>

In this section, we briefly introduce the spin dynamics of the planet (for a
more detailed treatment, see Paper I). We consider a star of mass $M_\star$
hosting an inner oblate planet of mass $m$ and radius $R$ on a circular orbit
with semi-major axis $a$ and an outer perturber of mass $m_{\rm p}$ on a
circular orbit with semi-major axis $a_{\rm p}$. We assume that the two orbits
are mutually inclined by the angle $I$. Denote $\bm{S}$ the spin angular
momentum and $\bm{L}$ the orbital angular momentum of the planet, and
$\bm{L}_{\rm p}$ the angular momentum of the perturber. The corresponding unit
vectors are $\uv{s} \equiv \bm{S} / S$, $\uv{l} \equiv \bm{L} / L$, and
$\uv{l}_{\rm p} \equiv \bm{L}_{\rm p} / L_{\rm p}$. The spin axis $\uv{s}$ of
the planet tends to precess around its orbital (angular momentum) axis $\uv{l}$,
driven by the gravitational torque from the host star acting on the planet's
rotational bulge. On the other hand, $\uv{l}$ and the disk axis $\uv{l}_{\rm p}$
precess around each other due to gravitational interactions. We assume $S \ll L
\ll L_{\rm p}$, so $\uv{l}_{\rm p}$ and $\uv{l}$ are nearly constant. The
equations of motion for $\uv{s}$ and $\uv{l}$ in this limit are
(\citealp{anderson2018teeter}, Paper I)
\begin{align}
    \rd{\uv{s}}{t}
        &= \omega_{\rm sl}\p{\uv{s} \cdot \uv{l}}\p{\uv{s} \times \uv{l}}
        \equiv \alpha\p{\uv{s} \cdot \uv{l}}\p{\uv{s} \times
        \uv{l}},\label{eq:dsdt1}\\
    \rd{\uv{l}}{t} &= \omega_{\rm lp}\p{\uv{l} \cdot \uv{l}_{\rm p}}\p{\uv{l}
        \times \uv{l}_{\rm p}} \equiv -g\p{\uv{l} \times \uv{l}_{\rm p}},
        \label{eq:dldt1}
\end{align}
where
\begin{align}
    \omega_{\rm sl} &\equiv \frac{3GJ_2 mR^2 M_\star}{2a^3 I\Omega_{\rm s}}
        = \frac{3k_q}{2k}\frac{M_\star}{m}\p{\frac{R}{a}}^3 \Omega_{\rm s},
            \label{eq:wsl}\\
    \omega_{\rm lp} &= \frac{3m_{\rm p}}{4M_\star}\p{\frac{a}{a_{\rm p}}}^3
        n.\label{eq:wlp}
\end{align}
In Eq.~\eqref{eq:wsl}, $\Omega_{\rm s}$ is the spin frequency of the inner
planet, $\mathcal{I} = k mR^2$ (with $k$ a constant) is its moment of inertia
and $J_2 = k_{\rm q}\Omega_{\rm s}^2 (R^3/Gm)$ (with $k_{q}$ a constant) is its
rotation-induced (dimensionless) quadrupole moment [for a body with uniform
density, $k=0.4, k_{\rm q}=0.5$; for the Earth, $k \simeq 0.331$ and $k_{\rm q}
\equiv k_2 / 3 \simeq 0.1$ \citep[e.g.][]{groten2004fundamental,
lainey2016quantification}]. In other studies, $3k_{\rm q} / 2 k$ is often
notated as $k_2 / 2C$ \citep[e.g.][]{millholland_disk}. In Eq.~\eqref{eq:wlp},
$n \equiv \sqrt{GM_\star/a^3}$ is the inner planet's orbital mean motion,  and
we have assumed $a_{\rm p}\gg a$ and included only the leading-order
(quadrupole) interaction between the inner planet and perturber. Following
standard notation, we define $\alpha = \omega_{\rm sl}$ and $g \equiv
-\omega_{\rm 1p} \cos I$ \citep[e.g.][]{colombo1966}.

@@

Using this result, we can then identify the long-term equilibria of the system
when $\Omega_{\rm s}$ is also evolving: these equilibria of the system must both
satisfy $\dot{\Omega}_{\rm s} = 0$ and be a CS that is stable in the presence of
tidal dissipation (i.e.\ satisfying $\rdil{\uv{s}}{t} = 0$).
Figure~\ref{fig:6equils006} describes the evolution of
Eqs.~(\ref{eq:dsdt_tide}--\ref{eq:dWsdt_tide}) qualitatively in the coordinates
$(\Omega_{\rm s}, \theta)$, along with the locations of CS1 and CS2. The circled
points in Fig.~\ref{fig:6equils006} satisfy the criteria to be long-term
equilibria, and we call them \emph{tidal Cassini Equilibria} (tCE). We number
tCE1 and tCE2 the tCE that are in CSs 1 and 2 respectively. The obliquities of
the tCE depend on the system architecture, which can be quantified using the
parameter
\begin{align}
    \eta_{\rm sync} &\equiv \big(\eta\big)_{\Omega_{\rm s} = n}
        = \eta \frac{\Omega_{\rm s}}{n}\nonumber\\
        &= \frac{1}{2}\frac{k}{k_{\rm q}}
            \frac{m_{\rm p}m}{M_\star^2}
            \p{\frac{a}{a_{\rm p}}}^3 \p{\frac{a}{R}}^3\cos(I).
            \label{eq:def_etasync}
\end{align}
In Fig.~\ref{fig:6equils006}, $\eta_{\rm sync} = 0.06$;
Figs.~\ref{fig:6equils050}--\ref{fig:6equils070} illustrate the change to the
tCE when $\eta_{\rm sync} = 0.5$ and $0.7$ respectively.

@@

    \includegraphics[width=\textwidth]{../initial/1_weaktide/6equils0_06.png}

@@

    used $\abs{g}t_{\rm s} = 10^2$ and evolved each example for $5t_{\rm s}$).
    The phase space evolution of the two thicker evolutionary trajectories (cyan
    and pink; those beginning at $\theta_{\rm i} = 120^\circ$) are shown in

@@

    }\label{fig:6equils006}

@@

    \includegraphics[width=\textwidth]{../initial/1_weaktide/6equils0_50.png}
    \caption{Same as Fig.~\ref{fig:6equils006} but for $\eta_{\rm sync} = 0.5$.
    The light-colored crosses and lines correspond to evolutionary trajectories
    using the same initial conditions as those shown in
    Fig.~\ref{fig:6equils006}.}\label{fig:6equils050}
    \includegraphics[width=\textwidth]{../initial/1_weaktide/6equils0_70.png}
    \caption{Same as Figs.~\ref{fig:6equils006} and~\ref{fig:6equils050} but for
    $\eta_{\rm sync} = 0.7$. Note that $\eta_{\rm sync} = 0.7 > \eta_{\rm c}
    \approx 0.574$ and tCE1 does not exist. The phase space evolution of the
    thick purple trajectory in the bottom panel (starting at $\theta_{\rm i}
    = 10^\circ$) is shown in Fig.~\ref{fig:trajs3}. }\label{fig:6equils070}

@@

    Fig.~\ref{fig:6equils006} colored in pink, for which $\eta_{\rm sync} =
    0.06$ and $I = 20^\circ$ (for which $\eta_{\rm c} = 0.574$). The initial
    conditions are $\Omega_{\rm s, i} = 2.5n$, $\theta_{\rm i} = 120^\circ$, and
    $\phi_{\rm i} = 0^\circ$, and we have used $\abs{g}t_{\rm s} = 10^2$ and
    have evolved the system for $5t_{\rm s}$. In the left-most panel, the
    trajectory's evolution in $\p{\Omega_{\rm s}, \theta}$ coordinates along
    with the curves indicating CS1, 2, and $\dot{\Omega}_{\rm s} = 0$ are
    re-displayed from Fig.~\ref{fig:6equils006}. The vertical dashed lines
    denote the values of $\Omega_{\rm s}$ for which the phase space evolution of
    the system is displayed in the right four panels. In each of these right
    four panels, the trajectory's evolution for a single circulation/libration
    cycle is displayed in $\p{\cos \theta, \phi}$ phase space for the labeled
    value of $\Omega_{\rm s}$. The system encounters the separatrix, undergoes a
    III $\to$ I transition, and converges to tCE1.

@@

    \caption{Same as Fig.~\ref{fig:trajs1} but for $\phi_{\rm i} = 286^\circ$,
    corresponding to the trajectory in Fig.~\ref{fig:6equils006} colored in
    cyan. The system encounters the separatrix, undergoes a III $\to$ II
    transition, and converges to tCE2. The small $\phi$ offset of tCE2 is
    visible (see Fig.~\ref{fig:mcs}). Note that the initial condition of this
    system and that displayed in Fig.~\ref{fig:trajs1} have the same initial
    $\theta_{\rm i}$ and $\Omega_{\rm s, i}$ and only different precessional
    phases $\phi_{\rm i}$.

@@

    \caption{Same as Fig.~\ref{fig:trajs1} but for $\eta_{\rm sync} = 0.7$ and
    $\theta_{\rm i} = 10^\circ$, corresponding to the purple trajectory shown in
    the bottom panel of Fig.~\ref{fig:6equils070}. Here, the system evolves
    along CS1 until the separatrix disappears, upon which it experiences large
    obliquity variations that damp due to tidal dissipation. The highly
    asymmetric shape in the third panel arises due to the strong tidal
    dissipation used in this simulation. The system finally converges to tCE2,
    the only tCE that exists for this value of $\eta_{\rm sync}$ (see bottom
    panel of Fig.~\ref{fig:6equils070}).}\label{fig:trajs3}

@@

There are two important conditions that can change the existence and stability
of the tCE\@. First, if $\eta_{\rm sync} > \eta_{\rm c}$, then tCE1 will not
exist, as can be seen in Fig.~\ref{fig:6equils070}\footnote{Strictly speaking,
$\eta_{\rm sync}$ can be slightly smaller than $\eta_{\rm c}$, as the planet's
spin is slightly subsynchronous at tCE1 when $\eta \approx \eta_{\rm c}$, e.g.\
see Fig.~\ref{fig:6equils050}.}. Secondly, tCE2 may not be stable if the tidal
phase shift is too large. Applying the results of Section~\ref{ss:tidal_equils},
we find that tCE2 is stable as long as $t_{\rm s} \geq t_{\rm s, c}$ where
\begin{equation}
    t_{\rm s, c} \equiv \frac{\sin \theta_{\rm tCE2}}{\abs{g} \sin
            I}\p{\frac{2n}{\Omega_{\rm s, tCE2}} - \cos \theta_{\rm tCE2}}.
\end{equation}
The tCE2 subscripts denote evaluation at tCE2. When $\eta_{\rm sync} \ll
\eta_{\rm c}$, we can use Eq.~\eqref{eq:def_tce2_approx} to further simplify:
\begin{equation}
    t_{\rm s, c}\approx \frac{1}{\abs{g}\sin I}\sqrt{\frac{2}{\eta_{\rm sync}
        \cos I}}. \label{eq:def_ts_crit}
\end{equation}
Note that this result differs in form from that given in
\citet{fabrycky_otides}, which assume that the orbital precession of $\uv{l}$ is
driven by the central star. Our expression is simpler and is applicable to
systems where the orbital precession of the inner planet is dominated by an
exterior planet.

@@

With this result, we can now consider the final fate of the inner planet's spin.
We begin by qualitatively examining the example trajectories shown in
Figs.~\ref{fig:6equils006}, for which we have integrated
Eqs.~(\ref{eq:dsdt_rot},~\ref{eq:dsdt_tide}--\ref{eq:dWsdt_tide}) and set $I =
20^\circ$. We discuss each of the six trajectories in turn:
\begin{itemize}
    \item The trajectory with initial conditions $\Omega_{\rm s, i} = 2.5n$ and
        $\theta_{\rm i} = 10^\circ$ (purple) has an initially prograde spin
        (i.e.\ in zone I, see Fig.~\ref{fig:1contours}) and directly evolves to
        tCE1 as the obliquity damps to zero.

@@

    \item We have chosen two trajectories, both with initial conditions
        $\Omega_{\rm s, i} = 2.5n$ and $\theta_{\rm i} = 120^\circ$ but with
        different initial precessional phases $\phi_{\rm i}$; we first discuss
        the pink trajectory, for which $\phi_{\rm i} = 0^\circ$. This trajectory
        originates in zone III, evolves towards the separatrix as tidal friction
        damps the obliquity, and crosses the resonance without being captured,
        upon which the obliquity continues to damp until the system converges to
        tCE1. The detailed phase space evolution of this trajectory is shown in
        Fig.~\ref{fig:trajs1}, where the outcome of the separatrix encounter is
        very visible.

    \item The other trajectory with initial conditions $\Omega_{\rm s, i} = 2.5n$
        and $\theta_{\rm i} = 120^\circ$ (light blue) only differs from the pink
        trajectory as its initial precessional phase is $\phi_{\rm i} \approx
        286^\circ$. Nevertheless, this trajectory originates in zone III,
        encounters the separatrix and is captured into the resonance (zone II),
        upon which tidal friction drives the system towards tCE2. The detailed
        phase space evolution of this trajectory is shown in
        Fig.~\ref{fig:trajs2}, where the resonance capture is displayed. Also
        visible in the final panel of Fig.~\ref{fig:trajs2} is the slight phase
        offset of CS2, i.e.\ $\phi_{\rm cs} < 180^\circ$, in agreement with the
        discussion in Section~\ref{ss:tidal_equils} (see Fig.~\ref{fig:mcs}).

@@

        A subtlety of initially subsynchronous spins can be seen here: note that
        the initial $\eta_{\rm i} = 0.6$ here, so the separatrix and CS1 do not
        exist initially ($\eta_{\rm c} = 0.574$). As such, naively, one expects
        initial convergence to CS2 and subsequent obliquity along with CS2 as
        the spin increases. However, due to the strong tidal dissipation and the
        proximity of $\eta_{\rm i}$ to $\eta_{\rm c}$, CS1 appears within a
        single circulation cycle, and the obliquity quickly damps to, and
        continues to evolve along, CS1.

    \item The trajectory with initial conditions $\Omega_{\rm s, i} = 0.1n$ and
        $\theta_{\rm i} = 100^\circ$ (teal) also has its obliquity rapidly
        damped to zero as it approaches spin-orbit synchronization and tCE1. For
        these initial conditions, we remark that this initial condition
        converges to tCE2 when a tidal dissipation rate of $\abs{g}t_{\rm s} =
        10^3$ is used, satisfying the intuitive analysis given in the previous
        paragraph.
\end{itemize}
For each of these six initial conditions, we also display their evolutionary
trajectories for $\eta_{\rm sync} = 0.5$ and $\eta_{\rm sync} = 0.7$ in
Figs.~\ref{fig:6equils050}--\ref{fig:6equils070}. The qualitative evolution of
these six examples only changes in a few important ways, so we will discuss just
a few points of interest:
\begin{itemize}
    \item For both $\eta_{\rm sync} = 0.5$ and $\eta_{\rm sync} = 0.7$, we see
        that both initial conditions with $\theta_{\rm i} = 120^\circ$ converge
        to tCE2. In fact, for these values of $\eta_{\rm sync}$, all initial
        conditions with $\theta_{\rm i} = 120^\circ$ will converge to tCE2
        regardless of $\phi_{\rm i}$.

    \item The two subsynchronous initial conditions both evolve to tCE2 for both
        $\eta_{\rm sync} = 0.5$ and $\eta_{\rm sync} = 0.7$, as in both cases
        $\eta_{\rm i} \gg \eta_{\rm c}$ and CS2 is the only low-obliquity spin
        equilibrium. The system then continues to evolve along CS2 until
        reaching a state satisfying $\dot{\Omega}_{\rm s} = 0$ as well.

    \item Of particular interest is the evolution starting from the initial
        conditions $\Omega_{\rm s, i} = 2.5n$ and $\theta_{\rm i} = 10^\circ$
        (purple) when $\eta_{\rm sync} = 0.7$. Figure~\ref{fig:trajs3} shows the
        detailed phase space evolution of this system, where it can be seen that
        the system initially evolves along the stable CS1, but is ejected when
        $\Omega_{\rm s}$ is sufficiently small that CS1 ceases to exist, upon
        which large obliquity variations eventually lead to convergence to tCE2,
        the only tCE that exists.
\end{itemize}
It can be seen that the dynamics are complex and can vary greatly depending on
the parameters used. In the case where the initial spin is subsynchronous, the
detailed outcome depends sensitively on $\eta_{\rm i}$ and the tidal dissipation
rate. For tractability, we restrict our subsequent discussion to the more
astrophysically common regime $\Omega_{\rm s, i} \gg n$, and we adopt the
fiducial value $\Omega_{\rm s, i} = 10n$.
